Multitask matrix completion for learning protein interactions across diseases

### 2016-09-01 paper.tex
new pdf

\begin{figure*}
\begin{floatrow}
\ffigbox[0.28\textwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1-eps-converted-to.pdf}
}
{
\caption{Multiple bipartite graphs: on the left are host proteins and on the right are proteins from different virus species. Edges represent protein interactions.}
\label{fig:hpppigraph}
}
\capbtabbox[0.68\textwidth]{%
\begin{footnotesize}
\begin{tabular}{c|r|r|r} 
\hline
Task $\rightarrow$ &{\textit{Influenza A} } &\textit{Hepatitis C} & {\textit{Ebola}}\\
& (28 strains) & (5 strains, many isolates)  & (3 strains) \\
number of HP PPIs (positives) & 848 & 981 & 90 \\\hline
number of non-homologous PPIs & 274 & 270 & 75 \\ \hline
\# of unique virus proteins in dataset & 54 & 151 & 2 \\\hline
\# of unique human proteins in dataset & 362 & 385 & 88 \\\hline
total \# of virus proteins in & 542 & 163 & 38 \\
UniProtKB across strains $^\dagger$ & & & \\ \hline
number of negatives & 84800 & 98100 & 9000 \\\hline
density of observed graph$^\ddagger$ (as \%) & .15 & .60 & .20 \\\hline
\multicolumn{4}{l}{\scriptsize{HP PPI: host-pathogen protein protein interactions}}\\
\multicolumn{4}{l}{\scriptsize{$\ddagger$ Density = (no. of positives + no. of negatives)/ total number of possible edges. The denominator was computed }} \\
\multicolumn{4}{l}{\scriptsize{\hspace{0.1cm} considering all proteins from all strains of the virus and $\approx$ 100,000 human proteins.}}\\
\multicolumn{4}{l}{\scriptsize{$\dagger$ this is the number of `reviewed' proteins on UniProtKB from all strains in our data}} \\ \hline
\end{tabular}
\end{footnotesize}
}
{
\caption{Tasks and their sizes (a column corresponds to one bipartite graph between human proteins and the virus indicated in the column header). More importantly, each task represents several strains of one virus species. We pool together PPI from all strains as the number of interactions per strain is very small and not representative. PPI data from different strains often involves homologous proteins (virus proteins or/and human proteins) as these were curated from different publications/studies involving the virus. The second row gives a conservative estimate of the number of non-homologous PPI. These were obtained by computing BLAST sequence similarity (see Section \ref{sec:phisto} for details). All pathogens are single stranded RNA viruses. The last row shows that each of our graphs is extremely sparse.} 
\label{datasets}
}
\end{floatrow}
\end{figure*}

@@

\begin{table*}[!b]\caption{Area Under the Precision-Recall curve (AUC-PR) on the PHISTO dataset. The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data. We report the average AUC-PR over 10 random train-test splits (stratified splits that maintain the class-skew of 1:100). The standard deviation is also shown. The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold. The Homolog baseline performs comparably to our BSL-MTL method on the \textit{Ebola} task. This is explained by the virus-human PPI graph, from several other viruses, that the Homolog baseline has access to.}
\label{resultsTable}
\begin{small}
\begin{center}
%\def\arraystretch{1.2}
\begin{tabular}{c|ccccccc}
\toprule
& \multicolumn{3}{c}{\textbf{10\% training, 90\% test}} && \multicolumn{3}{c}{\textbf{30\% training, 70\% test}} \\ \cline{2-4}\cline{6-8}
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} && \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
Homolog & \textbf{0.230}$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 && \textbf{0.311}$\pm$.03 & 0.180$\pm$.01 & 0.198$\pm$.01 \\
STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.130$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\ 
MMTL \citep{pontil04} & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\ 
%Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\ 
Sparse+low-rank \citep{chen2012} & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\ 
%Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
MTPL \citep{me_ismb_2013} & 0.217$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && 0.260$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
IMC \citep{nagarajan} & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
BSL-MTL (this work) & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
\end{tabular}
\end{center}
\end{small}
\end{table*}

@@

\begin{figure}
\centering
\includegraphics[scale=0.35, trim = 0 0.2cm 0 0.5cm]{fig2-eps-converted-to.pdf}
\caption{Principal component analysis (PCA) of virus proteins in the original feature space. The first two principal components are shown. Shape and color of the points indicates which virus that protein comes from.}%
\label{fig:pca1}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[scale=0.23, trim = 0 5.6cm 0 6cm]{pca_with_clusters.pdf}
\caption{Principal component analysis (PCA) of virus proteins in the projected subspace. The first two principal components are shown. Shape of the points indicates which virus that protein comes from. We have highlighted three clusters for which we show GO term enrichment analysis results in Table \ref{goterms}.}%
\label{fig:pca2}
\end{figure}

>>>>

\begin{figure*}
\capbtabbox[0.68\textwidth]{%
\begin{footnotesize}
\begin{tabular}{c|r|r|r} 
\hline
Task $\rightarrow$ &{\textit{Influenza A} } &\textit{Hepatitis C} & {\textit{Ebola}}\\
& (28 strains) & (5 strains, many isolates)  & (3 strains) \\
number of HP PPIs (positives) & 848 & 981 & 90 \\\hline
number of non-homologous PPIs & 274 & 270 & 75 \\ \hline
\# of unique virus proteins in dataset & 54 & 151 & 2 \\\hline
\# of unique human proteins in dataset & 362 & 385 & 88 \\\hline
total \# of virus proteins in & 542 & 163 & 38 \\
UniProtKB across strains $^\dagger$ & & & \\ \hline
number of negatives & 84800 & 98100 & 9000 \\\hline
density of observed graph$^\ddagger$ (as \%) & .15 & .60 & .20 \\\hline
\multicolumn{4}{l}{\scriptsize{HP PPI: host-pathogen protein protein interactions}}\\
\multicolumn{4}{l}{\scriptsize{$\ddagger$ Density = (no. of positives + no. of negatives)/ total number of possible edges. The denominator was computed }} \\
\multicolumn{4}{l}{\scriptsize{\hspace{0.1cm} considering all proteins from all strains of the virus and $\approx$ 100,000 human proteins.}}\\
\multicolumn{4}{l}{\scriptsize{$\dagger$ this is the number of `reviewed' proteins on UniProtKB from all strains in our data}} \\ \hline
\end{tabular}
\end{footnotesize}
}
{
\caption{Tasks and their sizes (a column corresponds to one bipartite graph between human proteins and the virus indicated in the column header). More importantly, each task represents several strains of one virus species. We pool together PPI from all strains as the number of interactions per strain is very small and not representative. PPI data from different strains often involves homologous proteins (virus proteins or/and human proteins) as these were curated from different publications/studies involving the virus. The second row gives a conservative estimate of the number of non-homologous PPI. These were obtained by computing BLAST sequence similarity (see Section \ref{sec:phisto} for details). All pathogens are single stranded RNA viruses. The last row shows that each of our graphs is extremely sparse.} 
\label{datasets}
}

\begin{floatrow}
\ffigbox[0.28\textwidth]{%
%trim option's parameter order: left bottom right top
%\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1-eps-converted-to.pdf}
\includegraphics[scale=0.35, trim = 0 0.2cm 0 0.5cm]{fig2-eps-converted-to.pdf}}
{
\caption{Principal component analysis (PCA) of virus proteins in the original feature space. The first two principal components are shown. Shape and color of the points indicates which virus that protein comes from.}%
\label{fig:pca1}
%\caption{Multiple bipartite graphs: on the left are host proteins and on the right are proteins from different virus species. Edges represent protein interactions.}
%\label{fig:hpppigraph}
}

\end{floatrow}
\end{figure*}

@@

\begin{figure*}
\capbtabbox[0.68\textwidth]{%
\begin{small}
\begin{center}
%\def\arraystretch{1.2}
\begin{tabular}{c|ccccccc}
\toprule
& \multicolumn{3}{c}{\textbf{10\% training, 90\% test}} && \multicolumn{3}{c}{\textbf{30\% training, 70\% test}} \\ \cline{2-4}\cline{6-8}
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} && \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
Homolog & \textbf{0.230}$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 && \textbf{0.311}$\pm$.03 & 0.180$\pm$.01 & 0.198$\pm$.01 \\
STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.130$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\ 
MMTL \citep{pontil04} & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\ 
%Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\ 
Sparse+low-rank \citep{chen2012} & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\ 
%Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
MTPL \citep{me_ismb_2013} & 0.217$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && 0.260$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
IMC \citep{nagarajan} & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
BSL-MTL (this work) & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
\end{tabular}
\end{center}
\end{small}
}
{
\caption{Area Under the Precision-Recall curve (AUC-PR) on the PHISTO dataset. The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data. We report the average AUC-PR over 10 random train-test splits (stratified splits that maintain the class-skew of 1:100). The standard deviation is also shown. The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold. The Homolog baseline performs comparably to our BSL-MTL method on the \textit{Ebola} task. This is explained by the virus-human PPI graph, from several other viruses, that the Homolog baseline has access to.}
\label{resultsTable}
}

\begin{floatrow}
\ffigbox[0.28\textwidth]{%
\includegraphics[scale=0.23, trim = 0 5.6cm 0 6cm]{pca_with_clusters.pdf}
}
{
\caption{Principal component analysis (PCA) of virus proteins in the projected subspace. The first two principal components are shown. Shape of the points indicates which virus that protein comes from. We have highlighted three clusters for which we show GO term enrichment analysis results in Table \ref{goterms}.}%
\label{fig:pca2}
}
\end{floatrow}
\end{figure*}

@@
